On The Impact of Atoms of Confusion in JavaScript Code

### 2021-04-26 4_Result.tex
update the survey + spell checker

\begin{table}[htbp]
\caption{Time in seconds to submit a correct answer}
\label{tab:difference-time-taken}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
Atom & Confusing Code & Clean Code & $\Delta$(\%) \\ \midrule
Comma Operator & 87.67 & 20.84 & -76.23 \\ 
Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
Post-Increment & 28.70 & 25.67 & -10.56 \\ 
Ommitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
Implict Predicate & 36.24 & 24.01 & -33.75 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
Ternary Operator & 41.80 & 39.34 & -5.90 \\ 
Pre-Increment & 30.71 & 42.45 & 38.19 \\  
Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
\end{tabular}
\end{scriptsize}
}
\end{table}

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hipothesis Testing (misunderstanding and cognitive workload)}
\label{tab:hypothesis-testing}

 \centering
 \begin{tabular}{lccc|cc}
   \toprule
Atom candidate & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
Ommitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
Implict Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
Ternary Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or  clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for
six atom candidates (Comma Operator, Automatic Semicolong Insertion,
Implicit Predicate, Logic
as Control Flow, Pre-Increment, and Arithmetic
as Logic). For the first four, the analysis suggests that the participants
need more time to predict the outcome of
the code snippet in the confusing code version. Interesting, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confusing version.
We also computed the effect size using the Cliff's Delta method. We found a large effect for the Comma Operator atom candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

@@

In this section we present the results of the 
interviews with the practitioners who agreed to participate
in the study. A total of 15 practitioners
took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion
about the \na atom candidates---by
(a) presenting them pairs of code snippets whose behavior were identical
and then (b) asking them to compare the code snippets readability.

@@

\subsubsection*{The Participants' perceptions of the atom candidates}
Supporting the results of the survey, Table~\ref{tab:interview-results1}
 shows that for eight out of the
nine scenarios surveyed, the majority of the respondents agree that the version of the
code without the atom of confusion is easier to understand, and in no case the \emph{neutral}
ratio was higher than the option for the version without atoms of confusion.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository. 
Only for the Ternary Operator atom candidate the participants
prefer the version of the code with the atom of confusion,
instead of the more cleaner version without the atom of confusion.
We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
Even for this atom candidate, some participants who preffered 
the \emph{left-hand-side} version (with the atom of confusion) 
still believed that the \emph{right-hand-side} version (without the atom of
confusion) was more readable.  
The following quotes were extracted from the transcripts with
three interviewees:

@@

The Pre-Increment atom candidate also caused such
divide in one of the interviewees, as they
regarded the \emph{\rhs version} (without atom of confusion)
as simpler to understand, but would still opt to write
code that contained the atom. In contrast to such opinion,
one of the participants found the version that contained the atom
of confusion more elegant, but recognized it was less readable,
and was willing to sacrifice elegance for readability.

@@

When analyzing the Logic as Control Flow atom candidate, one of
the interviewees gave an example of his own experience
that might motivate one avoid writing code using this construct:

@@

\begin{table}[!htb]
    \centering
    \caption{Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Arithmetic as Logic     & 0  & 93.33  & 6.67  \\
         Assignment as Value     & 20 & 60     & 20    \\
         Comma Operator          & 0  & 100    & 0     \\
         Ternary Operator        & 60 & 26.67  & 13.3  \\
         Implicit Predicate      & 20 & 73.33  & 6.67  \\
         Logic as Control Flow   & 20 & 60     & 20    \\
         Omitted Curly Braces    & 0  & 100    & 0     \\
         Post Increment          & 20 & 73.33  & 6.67  \\
         Pre-Increment           & 40 & 46.67  & 13.33 \\ \midrule
         \textsc{overall}        & 20 & 70.37  & 9.63  \\
         \bottomrule
    \end{tabular}
\end{table}

@@

We mined \minedprojects open source JavaScript repositories to understand how
often atoms of confusion arise in real software. 
Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigte
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atom Ternary Operator was found in all repositories we mined,
while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

Considering all JavaScript projects in our dataset, we found 
a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible for 87.97\% of this total: Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
account for \num{24558} (12.03\%) of the total number of occurrences, and the Comma Operator and Arithmetic as Logic are the atoms that arise less frequently (0.001\% in total with 232 and 165 occurrences, respectively).

Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
according to the thresholds of a previous
study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. As such, we found three
atoms as highly used (Ternary Operator, Omitted Curly Braces, and Post Increment),
four as commonly used (Pre-Increment, Assignment as Value,
Logic as Control Flow, and Automatic Semicolon Insertion),
and two atoms as little used (Arithmetic as Logic and Comma Operator).
We also calculated the frequency of each atom per KLOC and found the atom
Ternary Operator as the most occurring in practice (10.16 occurrences per KLOC).
%In our dataset of JavaScript projects, 
%We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}. %

The results of our survey and interviews suggest that the
Ternary Operator does not contribute significantly as a source of
misunderstanding (increasing the number of wrong answers in 9\% of the cases,
according to our survey). Nonetheless, the Post-Increment Expression
and Omitted Curly Braces atoms are listed in the top four sources of misunderstanding 
(see Table~\ref{tab:difference-correctness}). As such, fixing the
the Post-Increment and Omitted Curly Braces might solve most
of the unclear code caused by atoms of confusion in
JavaScript projects.

>>>>

\begin{table}[htbp]
\caption{Time in seconds to submit a correct answer}
\label{tab:difference-time-taken}
\centering{
\begin{scriptsize}
\begin{tabular}{lccc} \toprule
Atom & Confusing Code & Clean Code & $\Delta$(\%) \\ \midrule
Comma Operator & 87.67 & 20.84 & -76.23 \\ 
Automatic Semicolon Insertion & 46.08 & 22.04 & -52.17 \\ 
Post-Increment & 28.70 & 25.67 & -10.56 \\ 
Omitted Curly Braces & 48.85 & 30.00 & -38.58 \\ 
Assignment as Value & 52.47 & 48.95 & -6.71 \\ 
Implicit Predicate & 36.24 & 24.01 & -33.75 \\ 
Logic as Control Flow & 108.94 & 51.07 & -53.12 \\ 
Ternary Operator & 41.80 & 39.34 & -5.90 \\ 
Pre-Increment & 30.71 & 42.45 & 38.19 \\  
Arithmetic as Logic & 28.82 & 37.20 & 29.06 \\ \bottomrule
\end{tabular}
\end{scriptsize}
}
\end{table}

 % latex table generated in R 4.0.4 by xtable 1.8-4 package
 % Thu Apr 22 07:58:21 2021
 \begin{table*}[ht]
\caption{Hypotheses Testing (misunderstanding and cognitive workload)}
\label{tab:hypothesis-testing}

 \centering
 \begin{tabular}{lccc|cc}
   \toprule
Atom candidate & Chi-square test & Odds Ratio          & CI & Mann-Whitey U test & Cliff's Delta \\ 
     &         & (misunderstanding)  &                    & (workload)             & \\ \midrule
Comma Operator & 1.1727e-10 & 19.02 & (6.60, 68.22) & 5.8249e-08 & -0.53 \\ 
Automatic Semicolon Insertion & 5.8352e-11 & 39.33 & (9.21, 356.62) & 0.09802 & -0.16 \\ 
Post-Increment & 0.0015279 & 4.83 & (1.73, 15.74) & 0.12006 & 0.15 \\ 
Omitted Curly Braces & 0.050962 & 2.35 & (1.00, 5.77) & 0.92529 & -0.01 \\ 
Assignment as Value & 0.0034775 & 8.39 & (1.81, 79.12) & 0.82681 & 0.02 \\ 
Implicit Predicate & 0.01123 & 6.95 & (1.46, 66.56) & 0.0029039 & -0.29 \\ 
Logic as Control Flow & 0.292 & 1.54 & (0.73, 3.28) & 7.5862e-05 & -0.39 \\ 
Ternary Operator & 0.15896 & 2.73 & (0.74, 12.57) & 0.24073 & -0.12 \\ 
Pre-Increment & 0.70147 & 1.25 & (0.55, 2.85) & 0.0062981 & 0.27 \\ 
Arithmetic as Logic & 1 & 0.84 & (0.22, 3.12) & 0.01723 & 0.23 \\ 
    \bottomrule
 \end{tabular}
 \end{table*}

\subsubsection*{Hypothesis Testing}
We use the non-parametric \emph{Mann-Whitney U test}, also known as Wilcoxon rank-sum test, to
investigate the null hypothesis that developers 
spend the same amount of time to correctly
predict the outcome of a code snippet, regardless
of evaluating the confusing or  clean version
of the code. The results of Table~\ref{tab:difference-time-taken}
suggest that we should refute the null hypotheses for
six atom candidates (Comma Operator, Automatic Semicolon Insertion,
Implicit Predicate, Logic
as Control Flow, Pre-Increment, and Arithmetic
as Logic). For the first four, the analysis suggests that the participants
need more time to predict the outcome of
the code snippet in the confusing code version. Interesting, for the atom candidates Arithmetic as Logic and Pre-Increment, the participants take less time
to predict the outcome of the code snippets in
the confusing version.
We also computed the effect size using the Cliff's Delta method. We found a large effect for the Comma Operator atom candidate; and a moderate effect for the atom candidates
Logic as Control Flow, Implicit Predicate, Arithmetic
as Logic, and Pre-Increment.

@@

In this section we present the results of the 
interviews with the practitioners who agreed to participate
in the study. A total of 15 practitioners
took part of the interviews.
We collected information regarding programming
experience, familiarity with JavaScript, and their opinion
about the \na atom candidates we explored in our research.
To this end we first present them pairs of code snippets whose behavior were similar
and then ask the participants to discuss their preferences (i.e.,
code with or without atom).

@@

\subsubsection*{The Participants' perceptions of the atom candidates}
Supporting the results of the survey, Table~\ref{tab:interview-results1}
 shows that for eight out of the
10 scenarios surveyed, the majority of the respondents prefer the version of the
code without the atom of confusion, and in no case the \emph{neutral}
ratio was higher than the option for the version without atoms of confusion.
%% An example entry is contained in Appendix~\ref{}. 
%% For a full listing of the code snippets, visit the paper repository. 
Only for the Ternary Operator atom candidate the participants
prefer the version of the code with the atom of confusion,
instead of the version without the atom of confusion.
We show the code snippets for the Ternary Operator in Figure~\ref{code:ternary}.
Even for this atom candidate, some participants who preferred 
the \emph{left-hand-side} version (with the atom of confusion) 
still believed that the \emph{right-hand-side} version (without the atom of
confusion) was more readable.  
The following quotes were extracted from the transcripts with
three interviewees:

@@

The Pre-Increment atom candidate also caused such
divide in one of the interviewees, as they
regarded the \emph{\rhs version} (without atom of confusion)
as simpler to understand, but would still opt to write
code with the atom. In contrast to such opinion,
one of the participants found the version with the atom
of confusion more elegant, but recognized it was less readable,
and was willing to sacrifice elegance for readability.

@@

When analyzing the Logic as Control Flow atom candidate, one of
the interviewees gave an example of his own experience
that might motivate one to avoid writing code using this construct:

@@

\begin{table}[!htb]
    \centering
    \caption{Summary of participants' preferences for code snippets \textsc{with} and \textsc{without} atoms of confusion (\textsc{aoc}). Participants were also allowed to choose \textsc{neutral} when they thought both sides were equally readable.}
    \label{tab:interview-results1}
    \begin{tabular}{lrrr}\toprule
      & \multicolumn{3}{c}{\textsc{Preference (\%)}} \\
      \cmidrule(lr){2-4}
         \textsc{atom}           & \multicolumn{1}{c}{\textsc{with aoc}}
                                      &  \multicolumn{1}{c}{\textsc{without aoc}}
                                               & \multicolumn{1}{c}{\textsc{neutral}} \\ \midrule
         Comma Operator                  & 0  & 100    & 0     \\
         Automatic Semicolon Insertion  & 0  & 80     & 0     \\
         Post-Increment                  & 20 & 73.33  & 6.67  \\
         Omitted Curly Braces            & 0  & 100    & 0     \\
         Assignment as Value             & 20 & 60     & 20    \\
         Implicit Predicate              & 20 & 73.33  & 6.67  \\
         Logic as Control Flow           & 20 & 60     & 20    \\
         Ternary Operator                & 60 & 26.67  & 13.3  \\
         Pre-Increment                   & 40 & 46.67  & 13.33 \\ 
         Arithmetic as Logic             & 0  & 93.33  & 6.67  \\ \ midrule
         \textsc{overall}                & 18 & 71.33  & 8.64  \\
         \bottomrule
    \end{tabular}
\end{table}

@@

We mined \minedprojects open source JavaScript repositories to understand how
often atoms of confusion arise in real software. 
Similarly to previous studies~\cite{DBLP:conf/msr/GopsteinZFC18,DBLP:journals/ese/MedeirosLAAKRG19}, which investigate
the prevalence of atoms of confusion in open source \clang and \cpplang projects, we found that atoms of confusion
frequently arise in JavaScript open source systems. For instance, the six most frequently found atoms
occur in at least 80\% of the projects. Considering the extremes, atom Ternary Operator was found in all repositories we mined,
while atom Comma Operator occurred in only 15\% of them (as seen in Table~\ref{tab:occurrences-summary}).

@@

Considering all JavaScript projects in our dataset, we found 
a total of {\color{red}\num{364873}} atoms of confusion, though three atoms are responsible
for {\color{red}87.97\% of this total}: Ternary Operator, Omitted Curly Braces, and the Post Increment. The remaining atoms
account for \num{24558} (12.03\%) of the total number of occurrences, and the
Comma Operator and Arithmetic as Logic are the atoms that arise less
frequently (0.001\% in total with 232 and 165 occurrences, respectively).

Regarding frequency, we proceeded with a classification of \emph{atom occurrence}
according to the thresholds of a previous
study~\cite{DBLP:journals/ese/MedeirosLAAKRG19}. As such, we found four
atoms as highly used (Implicit Predicate, Ternary Operator, Omitted Curly Braces, and Post Increment),
four as commonly used (Pre-Increment, Assignment as Value,
Logic as Control Flow, and Automatic Semicolon Insertion),
and two atoms as little used (Arithmetic as Logic and Comma Operator).
We also calculated the frequency of each atom per KLOC and found the atoms
Implicit Predicate (19.89 occurrences per KLOC) and Ternary Operator (10.16
occurrences per KLOC) as the most occurring
in practice.
%In our dataset of JavaScript projects, 
%We found the Ternary Operator atom occurring more frequently than the Omitted Curly Braces, differently from what has been reported in a previous work~\cite{DBLP:conf/msr/GopsteinZFC18}. %

The results of our survey and interviews suggest that the
Ternary Operator does not contribute significantly as a source of
misunderstanding (increasing the number of wrong answers in 9\% of the cases,
according to our survey). Nonetheless, the Post-Increment Expression
and Omitted Curly Braces atoms are listed in the top four sources of misunderstanding 
(see Table~\ref{tab:difference-correctness}), and also frequently appear
in open source systems. As such, fixing the
the Post-Increment and Omitted Curly Braces might solve a considerable
amount unclear code caused by atoms of confusion in
JavaScript projects.
